Hierarchical Inference of the Intrinsic Ellipticity Distribution of Galaxies during Weak Lensing Analysis

### 2013-04-15 docs/ms.tex
add fig

\subsection{Posterior Samples}
The probability distribution can be visualised by plotting points at each sample
position. An example in the x, y plane is shown for the posterior distribution of
$g$ in 64 sky patches in Figure 1. There are 8 galaxies in each patch in the simulation.
The number density of samples is proportional to the probability density. The figure shows that the two parameters of $\vec{g}$
do not have strong correlation, since the shape of smaples is close to a circle.
We also shows the histogram of g1 in a single sky patch in Figure 2. Not surprisingly it is close to a Gaussian distribution.
If the probability distribution of parameters are perfect Gaussian, the most probable estimates and uncertainty 
is just the mean and standard deviation of the Gaussian distribution. However, in pratical, we find that the distribution 
of reduced shear $\vec{g}$ will  get closer to Gaussian, when larger number of galaxies in one patch  is used. When the number of galaxies in 
one patch is equal to or larger than 8, we found it's eligible to make such an approximation that to use sample mean and standard 
deviation as the most probable estimates and uncertainty. We'll justify this approximation with a bias test in the following
sub section.

@@

@@

As described in section 3.1, we use the mean of posterior samples as the most probable $g_{i}$ and 
the standard deviation as uncertainty. A weighted minimum $\chi^{2}$ linear fit is performed to get the
most-likely value and uncertainty of m and c. for $g_{1}$ component we get $m=-6.09\times10^{-4}\pm2.39\times10^{-3}$
and $c=(1.82\pm1.44)\times10^{-4}$, while for $g_{2}$ component
we have $m=-3.48\times10^{-4}\pm2.14\times10^{-3}$ and $c=(2.01\pm1.29)\times10^{-4}$.
Due to the limitation of galaxy number we are not able to get an exact value for m and c. However in our result 
both the most-like and uncertainty of m are around $10^{-3}$ and for c they are around $10^{-4}$ . Due to the
relatively large uncertainty, it is only an rough estimate of bias level.
\begin{figure}
\includegraphics[scale=0.5]{fig/g1bias}

\caption{The figure shows the estimated $g_{1}$ values minus the true (input)
$g_{1}^{true}$. The distribution range of $g_{1}^{true}$ is from
-0.1 to 0.1. Originally there are 640,000 true $g_{1}$. We bin the
plot for presentation purpose. The data point and error bars are the weighted
 average and  variance of $\delta g_{1}$ in each bin. The blue
line is the result of weighted linear fit of the original 640,000 data points
with sample standard deviation as error bar.}
\end{figure}
Suppose we have perfect measurements on galaxy shape. The measured ellipticity then equals to the true lensed ellipticity. With equation (1) the mean ellipticity
estimator has following expression :
\begin{equation}
<\epsilon_{\ell}>=g^{t}+<\epsilon_{0}>+<(\epsilon_{0}+g^{t})[-g^{t*}\epsilon_{0}+O(g^{t*}\epsilon_{0})]>
\end{equation}
When the unlensd ellipticity is isotropic ,we have $<\epsilon_{\ell}>=g^{t}$ . In this case, the mean ellipticity estimator is 
unbiased.
Meanwhile we find that simply using the posterior mean as the estimator of $g$ will lead to a significant bias. 
Figure 3 shows the result of bias test for both mean ellipticity and posterior mean estimator.

@@

In this section we calculate the mean square error of mean of posterior samples from the true(input) $g$.  We show that the mean of posterior samples has smaller mean square error than the mean ellipticity estimator but has higher mean square error than the direct Bayesian estimates which could be made if we knew the exact intrinsic shape distribution.
In figure 5 we show how the mean square error changes with the number of galaxies and patches, and make a comparison with mean ellipticity estimator and direct Bayesian estimates. From the figure we can see that when the number of sky patches is small our hierarchical Bayesian method does not necessarily has lower mean square error than mean ellipticity estimator. However it becomes smaller and approach to the result of direct Bayesian estimate which is the best estimate we can possibly make.

@@

\subsection{Measurement Error}
In this section we  discuss how the shape measurement error 
affects the variance and bias of shear inference.
In principle, if the likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$
is known the likelihood $p(\epsilon_{ob}|g,\alpha)$ can be calculated using the convolution as equation (7).
However, due to the limitation of computational time we avoid to do this convolution but use $\epsilon_{ob}$
as an estimator of $\epsilon_{\ell}$.
The likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$ we use in the simulation is a 2D
Gaussian with mean equals $\epsilon_{\ell}$ and $\sigma$ equals 0.05 .
Figure 5 shows the mean square error of the sample mean estimates in comparison with mean ellipticity estimator.
\begin{figure}
\includegraphics[scale=0.48]{fig/errorN8}
\caption{Figure 5 shows how the mean square error of sample mean estimate varies with the number of patches used, while shape measurement error is considered as described in Section 3.4 .
The horizontal line denotes the mean square of mean ellipticity estiimator which does not depend on number of patches. }
\end{figure}
We can see that our proposed method still works well under shape measurement error, even when we do not have the exact likelihood
of true lensed ellipticity. It means that this hierarchical Bayesian method does not only naturally cope with model fitting shape measurement 
method but is also compatible with ellipticity point estimators.

>>>>

\subsection{Posterior Samples}
The probability distribution can be visualised by plotting points at each sample
position. An example in the x, y plane is shown for the posterior distribution of
$g$ in 64 sky patches in Figure 1. There are 8 galaxies in each patch in the simulation.
The number density of samples is proportional to the probability density. The figure shows that the two parameters of $\vec{g}$
do not have strong correlation, since the shape of smaples is close to a circle.
We also shows the histogram of g1 in a single sky patch in Figure 3. Not surprisingly it is close to a Gaussian distribution.
If the probability distribution of parameters are perfect Gaussian, the most probable estimates and uncertainty 
is just the mean and standard deviation of the Gaussian distribution. However, in pratical, we find that the distribution 
of reduced shear $\vec{g}$ will  get closer to Gaussian, when larger number of galaxies in one patch  is used. When the number of galaxies in 
one patch is equal to or larger than 8, we found it's eligible to make such an approximation that to use sample mean and standard 
deviation as the most probable estimates and uncertainty. We'll justify this approximation with a bias test in the following
sub section.

@@

\begin{figure}
\end{figure}
\begin{figure}
\includegraphics[scale=0.3]{fig/alpha_beta}
\includegraphics[scale=0.3]{fig/stp10}

\caption{The upper panel show the MCMC samples of shape parameters of the Beta function model. The true(input) value is (2.8,2.8) as indicated by the two straight lines in the graphic. The lower panel shows the sample mean of the shape parameters of the step function model. The curve denotes the true shape distribution $P(|\epsilon_{0}|)$, which is Beta (2.8,2.8) in this case.}
\end{figure}

@@

As described in section 3.1, we use the mean of posterior samples as the most probable $g_{i}$ and 
the standard deviation as uncertainty. A weighted minimum $\chi^{2}$ linear fit is performed to get the
most-likely value and uncertainty of m and c. For $g_{1}$ component we get $m=-6.09\times10^{-4}\pm2.39\times10^{-3}$
and $c=(1.82\pm1.44)\times10^{-4}$, while for $g_{2}$ component
we have $m=-3.48\times10^{-4}\pm2.14\times10^{-3}$ and $c=(2.01\pm1.29)\times10^{-4}$. Figure 3 shows the result of bais test for the $g_{1}$ component.
Due to the limitation of galaxy number we are not able to get an exact value for m and c. However in our result 
both the most-like and uncertainty of m are around $10^{-3}$ and for c they are around $10^{-4}$ . Due to the
relatively large uncertainty, it is only an rough estimate of bias level.
\begin{figure}
\includegraphics[scale=0.5]{fig/g1bias}

\caption{The figure shows the estimated $g_{1}$ values minus the true (input)
$g_{1}^{true}$. The distribution range of $g_{1}^{true}$ is from
-0.1 to 0.1. Originally there are 640,000 true $g_{1}$. We bin the
plot for presentation purpose. The data point and error bars are the weighted
 average and  variance of $\delta g_{1}$ in each bin. The blue
line is the result of weighted linear fit of the original 640,000 data points
with sample standard deviation as error bar.}
\end{figure}
Suppose we have perfect measurements on galaxy shape. The measured ellipticity then equals to the true lensed ellipticity. With equation (1) the mean ellipticity
estimator has following expression :
\begin{equation}
<\epsilon_{\ell}>=g^{t}+<\epsilon_{0}>+<(\epsilon_{0}+g^{t})[-g^{t*}\epsilon_{0}+O(g^{t*}\epsilon_{0})]>
\end{equation}
When the unlensd ellipticity is isotropic ,we have $<\epsilon_{\ell}>=g^{t}$ . In this case, the mean ellipticity estimator is 
unbiased.
Meanwhile we find that simply using the posterior mean as the estimator of $g$ will lead to a significant bias. 
Figure 5 shows the result of bias test for both mean ellipticity and posterior mean estimator.

@@

In this section we calculate the mean square error of mean of posterior samples from the true(input) $g$.  We show that the mean of posterior samples has smaller mean square error than the mean ellipticity estimator but has higher mean square error than the direct Bayesian estimates which could be made if we knew the exact intrinsic shape distribution.
In figure 6 we show how the mean square error changes with the number of galaxies and patches, and make a comparison with mean ellipticity estimator and direct Bayesian estimates. From the figure we can see that when the number of sky patches is small our hierarchical Bayesian method does not necessarily has lower mean square error than mean ellipticity estimator. However it becomes smaller and approach to the result of direct Bayesian estimate which is the best estimate we can possibly make.

@@

\subsection{Measurement Error}
In this section we  discuss how the shape measurement error 
affects the variance and bias of shear inference.
In principle, if the likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$
is known the likelihood $p(\epsilon_{ob}|g,\alpha)$ can be calculated using the convolution as equation (7).
However, due to the limitation of computational time we avoid to do this convolution but use $\epsilon_{ob}$
as an estimator of $\epsilon_{\ell}$.
The likelihood of true lensed ellipticity $p(\epsilon_{ob}|\epsilon_{\ell})$ we use in the simulation is a 2D
Gaussian with mean equals $\epsilon_{\ell}$ and $\sigma$ equals 0.05 .
Figure 7 shows the mean square error of the sample mean estimates in comparison with mean ellipticity estimator.
\begin{figure}
\includegraphics[scale=0.48]{fig/errorN8}
\caption{Figure 5 shows how the mean square error of sample mean estimate varies with the number of patches used, while shape measurement error is considered as described in Section 3.4 .
The horizontal line denotes the mean square of mean ellipticity estiimator which does not depend on number of patches. }
\end{figure}
We can see that our proposed method still works well under shape measurement error, even when we do not have the exact likelihood
of true lensed ellipticity. It means that this hierarchical Bayesian method does not only naturally cope with model fitting shape measurement 
method but is also compatible with ellipticity point estimators.
